Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-03-19 paper.qmd
use new figure and add formula

```{r}
library(tidyverse)
library(gt)
library(gtExtras)

tribble(
  ~resolution, ~hex_area, ~hex_edge,
  2L , 86801.780398997, 182.5129565,
  3L,  12393.434655088,  68.97922179,
  4L,   1770.347654491,  26.07175968

) |> 
  gt() |> 
   cols_label(
    resolution ~ "Resolution",
    tidyselect::ends_with("area") ~ "Cell Area {{*persons* km^-2}}",
    # hex_area ~ md("Cell Area ({km^2}})"),
    hex_edge ~ "Edge length ({km}})"
  ) |> 
    gt::tab_header(
    title = md("H3 Cell Statistics"),
    subtitle = md("average hexagons")
  ) |> 
  opt_align_table_header(align = "left")
```

@@

@fig-density-winter-res-2 provides a promising result in the hourly flight
density calculation which at least shows
expected outcomes such as cells at zero density for war or no-fly zones such as
Ukraine, Libia or Syria.

::: {#fig-density-winter-res-2}

```{r density-winter-res-2}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("media/figures/flight_density_2024-08-01T12_2.png")
```

Flight density on 2024-08-01T12 for H3 cells at resolution 2 covering EUROCONTROL Area.
:::

@@

From Equation 15 in [@patera2008], the impact (on the ground)
probability density, $\sigma(L, i)$, at latitude $L$ of a reentring object
whose orbit has inclination $i$ is given by:

@@

>>>>

```{r}
#| label: tbl-resolutions
#| tbl-cap: H3 hexagons, average values
#| tbl-cap-location: bottom

library(tidyverse)
library(gt)
library(gtExtras)

tribble(
  ~resolution, ~hex_area, ~hex_edge,
  2L , 86801.780398997, 182.5129565,
  3L,  12393.434655088,  68.97922179,
  4L,   1770.347654491,  26.07175968
) |> 
  gt() |> 
   cols_label(
    resolution ~ "Resolution",
    hex_area ~ md("Cell Area"),
    hex_edge ~ md("Edge length")
  ) |> 
    gt::tab_header(
    title = md("H3 Cell Statistics"),
    subtitle = md("average hexagons")
  ) |> 
  cols_units(
    hex_area = "km^2",
    hex_edge = "km"
  ) |>
  opt_align_table_header(align = "left")
```

@@

@fig-density-winter-res-3 provides a promising result in the hourly flight
density calculation which at least shows
expected outcomes such as cells at zero density for war or no-fly zones such as
Ukraine, Libia or Syria.

::: {#fig-density-winter-res-3}

```{r density-winter-res-3}
#| out-width: "110%"
#| echo: false
knitr::include_graphics("media/figures/flight_density_2024-08-01T12_3.png")
```

Flight density in the EUROCONTROL Area on 2024-08-01T12 (H3 cells at resolution 3).
:::

@@

## Impact probability density

From Equation 15 in [@patera2008], the (on the ground) impact probability
density, $\sigma(L, i)$, at latitude $L$ of a reentring object
whose orbit has inclination $i$ (see @fig-orbital-inclination) is given by:

@@

::: {#fig-orbital-inclination}

```{r orbital-inclination}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("media/figures/orbital_inclination_clip.png")
```

Orbital inclination is the angle between the plane of an orbit and the equator. An orbital inclination of 0° is directly above the equator, 90° crosses right above the pole, and 180° orbits above the equator in the opposite direction of Earth’s spin. (NASA illustration by Robert Simmon.).
:::

## Aircraft effective exposed area
From Patera

> Even for a ballistic coefficient of 50 lbs/ft2, the flight path angle
> of a reentering object at 40,000 feet is -86.6 degrees, which corresponds
> to nearly vertical descent.
> Since most debris objects have ballistic coefficients equal to or 
> less than 50 lbs/ft2, the vertical descent assumption is a very
> accurate approximation.

$$
\varepsilon_i = \frac{s_{a} a_{h} + s_{d} a_{v}}{s_d}
$$
where $s_{a}$ is the average aircraft speed, $s_{d}$ is the average debris 
object fall speed, $a_v$ area of the aircraft as viewed from above,
$a_h$ area of the aircraft as viewed from the from the front of the aircraft.
